\sffamily \Large {\bfseries Introspection dynamics in asymmetric multiplayers games}

### 2023-03-01 latex/article.tex
Revised abstract, intro and discussion by Sap

\begin{document}
\maketitle
\onehalfspacing
\section*{Abstract}
\section*{Abstract}

@@

Most of these situations cannot be captured by the sum of several pairwise interactions. Thus, to account for such non-linearities, one needs to consider multiplayer games \cite{Gokhale:DGAA:2014}. A well-known effect that only emerges when more than two players are present is the ``second-order free-riding problem" \cite{Fowler:PNAS:2005}. A natural solution to maintain pro-social behavior in a community is to monitor and punish defectors (and/or reward cooperators). However, most forms of sanctioning are considerably costly \cite{Henrich:Science:2006}. Therefore, an additional (second-order) dilemma is at stake: deterring defection is beneficial for all, but an individual that does not pay the associated cost has an advantage over the ones that do. As long as cooperation is incentivized by punishment (or rewards) of others, individuals can take advantage by defecting in this second-order dilemma. As such, this tempts everyone not to engage in costly punishment (or rewarding), and cooperation can break down. For this reason, peer-punishment and sanctioning institutions can be hard to sustain without additional mechanisms \cite{Panchanathan:Nature:2004, Perc:SciRep:2012, Hilbe:SciRep:2012, Couto:JTB:2020, Pal:NatCom:2022}.

@@

Another interesting effect that can be explored with multiplayer games is the scale or size of the interaction itself. In situations that require some sort of coordination and where expectations on others play an important role in one's decisions, a growing group size might hinder the optimal outcome \cite{Skyrms:book:2003}. Likewise, it has been shown that it is hard to cooperate in large groups \cite{Santos:PNAS:2011, Hilbe:JTB:2015}.
This is not a general effect, though. It can happen that adding more players actually promotes cooperation. Gokhale and Traulsen (2011) investigate the 2-player and 3-player cases of a task allocation problem \cite{gokhale:JTB:2011}. Although the underlying rules defining the interactions are the same for the two cases, simply introducing a third player changes the overall game, resulting in a decrease in free-riding.

@@

Additionally, interaction group size can vary in a population of players. There, not only the average group size can have an important effect, but also the variance of the group size distribution \cite{Pena:Evolution:2011, Broom:BMB:2019}.

@@

\noindent Complexity further increases when players differ significantly among themselves. This diversity can be captured by asymmetric games \cite{Taylor:JAP:1979, Schuster:AB:1981, Gaunersdorfer:TPB:1991, Hofbauer:JMB:1996, Hofbauer:GEB:2005, Ohtsuki:JTB:2010, McAvoy:PlosCB:2015, Veller:JET:2016, Hauser:Nature:2019}. In symmetric games, all players are indistinguishable. Thus, to fully characterise the state of the game, we only require to know the number of players playing each strategy. Conversely, in asymmetric games, players can differ in their available actions and in their incentives to choose each action; players are of different types or roles. Therefore, they can have uneven effects on others' payoffs too. For example, in public goods games and collective-risk dilemmas, players can have different initial endowments (or wealth), productivitites, costs, risk perceptions, or risk exposures \cite{Milinski:CC:2011, Vasconcelos:PNAS:2014, Abouchakra:JTB:2014, Hauser:Nature:2019, Merhej:JAIR:2022, Wang:PTRSB:2023}. Hence, to fully describe the state of the game, we need to know which of the players is doing what. This greatly increases the size of the game's state space; even more so, for more than two players.  \\ \\ 

@@

% deterministic EGT
\noindent From evolutionary game theory (EGT) \cite{Maynard-Smith:Nature:1973, Maynard-Smith:book:1982, Hofbauer:book:1998, Nowak:book:2006} to learning models \cite{Sandholm:BioSys:1996, Fudenberg:book:1998b, Macy:PNAS:2002, Pangallo:GEB:2022}, game theoretic frameworks have been widely used to study strategic behavior. %, in particular, in multiplayer and asymmetric games.
%In evolutionary game theory, we assume a population of players and a mechanism by which strategies evolve over time. More successful strategies tend to persist in the population, either by genetic or cultural inheritance. The first key concept in evolutionary game theory is ``evolutionary stable strategy" (ESS),  introduced in 1973, in the seminal paper of John Maynard Smith and George Price \cite{Maynard-Smith:Nature:1973}. A (monomorphic) population playing an ESS cannot be invaded by mutants. 
\noindent The concept of ``evolutionary stable strategy" (ESS), originally proposed for parwise encounters \cite{Maynard-Smith:Nature:1973}, was extended for multiplayer games \cite{Palm:JMB:1984, Broom:BMB:1997, Bukowski:IJGT:2004}. Also the well-known replicator equation \cite{Taylor:MB:1978, Hofbauer:book:1998} easily comprises multiplayer games \cite{Hauert:JTB:2006a, gokhale:PNAS:2010, Pena:Evolution:2011, Cressman:PNAS:2014, Pena:JTB:2014}. More recently, the replicator-mutator equation was applied to study the dynamics of multiplayer games, too \cite{Duong:DGAA:2020}. 
As for asymmetric games, a few additional assumptions are needed. For example, if there are two different types of players, typically, either there are two populations co-evolving (``bimatrix games" \cite{Hofbauer:book:1998, Gokhale:PRSB:2012, Tuyls:SciRep:2018}) or there is a single population of players where each can play the two types or roles (``role games") \cite{Hofbauer:book:1998}. For role games, a strategy must include the action played in each of the two roles. The case of asymmetric games with more than two players is substantially less studied within deterministic EGT. Gokhale and Traulsen (2012) and Zhang et al. (2022) are two exceptions \cite{Gokhale:PRSB:2012, Zhang:arxiv:2022}. Notably, although these works study multiplayer games, they consider, at most, two different types (drawn from two populations), which leaves out the exploration of full asymmetry.
% models of multiple populations seem unnatural
\\ \\
% stochastic EGT
\noindent Also stochastic evolutionary game dynamics \cite{Nowak:Nature:2004, Traulsen:bookchapter:2009} provides several models for studying multiplayer and asymmetric games. 
%Here, the quantities of interest are fixation probabilities and fixation times (i.e., the probability and the average time that a single mutant overtakes an initially monomorphic population) for when mutations are rare \cite{Fudenberg:TPB:2006}, or the average strategy abundance (i.e., the stationary distribution of strategy frequencies) for when mutation rates take intermediate values \cite{antal:JTB:2009a, antal:JTB:2009b}. 
%These frameworks have been applied to asymmetric games assuming co-evolving (finite) subpopulations of different player types. 
Fixation probabilities \cite{Fudenberg:TPB:2006} for asymmetric 2-player \cite{Sekiguchi:DGA:2017}, asymmetric 3-player games \cite{Sekiguchi:DGAA:2022}, and symmetric multiplayer games \cite{Kurokawa:PRSB:2009, gokhale:PNAS:2010} were recently derived. Furthermore, average strategy abundances \cite{antal:JTB:2009a, antal:JTB:2009b} were obtained only for 2-player asymmetric games \cite{Ohtsuki:JTB:2010, Sekiguchi:PA:2013} or multiplayer symmetric games \cite{gokhale:JTB:2011, Wu:Games:2013, Kroumi:JMB:2022}. For a review on evolutionary multiplayer games both in infinitely large populations as well as in finite populations, we refer to Gokhale and Traulsen (2014) \cite{Gokhale:DGAA:2014}.\\

@@

%[Learning models]
%learning models in general
\noindent Learning models (of strategic behavior) take a different approach from EGT \cite{Sandholm:BioSys:1996, Fudenberg:book:1998b, Macy:PNAS:2002, Hofbauer:GEB:2005, Tuyls:bookchapter:2005, Galla:PNAS:2013, Barfuss:PRE:2019, Barfuss:PNAS:2020, Pangallo:GEB:2022}. There is no evolution of strategies in a population necessarily, but a process by which individuals learn strategies dynamically. 
%Despite the conceptual disparities, there have been already some efforts in unifying the two lines of research of evolutionary game theory and multiagent reinforcement learning \cite{Macy:PNAS:2002, Tuyls:bookchapter:2005, Bloembergen:JAIR:2015, Zhang:arxiv:2022}. 
% Clearly, we can distinguish two different pathways that led to ``learning in games": one which stemmed from the broad field of reinforcement learning \cite{Sandholm:BioSys:1996} and the other from the classical game theory itself \cite{Fudenberg:book:1998b}. The ``theory of learning in games" was proposed as a refinement of classical equilibrium concepts, like Nash equilibrium, where strict assumptions of common knowledge of rationality are made. Instead, learning consists of ``\textit{an alternative explanation that equilibrium arises as the long-run outcome of a process in which less than fully rational players grope for optimality over time}" \cite{Fudenberg:book:1998b}. In this context, several learning models were introduced, such as fictitious play \cite{Gaunersdorfer:GEB:1995} or perturbed best response dynamics \cite{Hofbauer:GEB:2005}.
%learning models for asymmetric and multiplayer games? fictious play for multiplayer games \cite{fudenberg:book:1998b} 
%can we say most of learning models have no analytical or exact results?
\noindent Introspection dynamics has recently proven to be a useful learning model for tackling asymmetric games \cite{Couto:NJP:2022, Hauser:Nature:2019, McAvoy:PNASnexus:2022, Schmid:PlosCB:2022}. 
This feature comes from the fact that players update their strategies by exploring their own set of strategies in the following simple way: each time, after a round of the game, a random player considers a random alternative strategy; they compare the payoff that it would have given them to their current payoff; if the new strategy would provide a higher payoff, it is more likely adopted on the next round. We describe the model formally in the next section. 
While in Couto et al. (2022) \cite{Couto:NJP:2022} only 2-player games were considered, this framework is general enough to account for multiple players.  Particularly, introspection dynamics allows a natural exploration of full asymmetry in many-player games compared to population models. For example, in imitation dynamics, one needs to specify who imitates who \cite{Vasconcelos:PNAS:2014}, as when players differ much, it might not make sense to assume they imitate any other player with the same likelihood. Introspection avoids this assumption because players' decisions only depend on their own payoffs.
\\ \\
%[Our proposal]
\noindent Here, we extend previous results of pairwise games under introspection dynamics to multiplayer games. 
First, we derive a formula that allows us to numerically compute the stationary distribution of introspection dynamics for any multiplayer asymmetric game. 
Second, we obtain explicit expressions of the stationary distribution for two special cases. These cases are additive games (where the payoff difference that a player gains by unilaterally switching to a different action is independent of the actions of their co-players), and symmetric multiplayer games with two strategies. 
To illustrate our theoretical results, we analyse various multiplayer asymmetric social dilemmas, extending the framework in \cite{Hauert:JTB:2006a} to asymmetric games. Finally, we also study the asymmetric version of a public goods game with a rewarding stage \cite{Pal:NatCom:2022}.

@@

\noindent We use the Eq. (\ref{Eq:stationary_dist_lpgg}) to analyze the stationary behaviour of players after a long run of introspection in a linear public goods game (LPGG). First, we study the simplest case where all players are symmetric (the cost and benefit for all the 4 players are $c$ and $b$). Since all players are identical, the states of the game can be enumerated by counting the number of cooperators in the state. There can only be 5 distinct states of the game (from 0 to 4 cooperators).

@@

\noindent Naturally, $\beta$ plays an important role in determining the overall cooperation in the long run. When $\beta$ is low, average cooperation varies weakly with the strength of the dilemma, $b/N - c$ (Fig. \ref{Fig:LPGG-symmetric}c). Even when the temptation to defect is high ($b/N - c = -2$), players cooperate with a non-zero probability. Similarly, when cooperation is highly beneficial and strictly dominates defection ($b/N - c = 2$), players defect sometimes. At higher values of $\beta$, the stationary behaviour of players is more responsive to the payoffs and thus reflects an abrupt change near the parameters where the game transitions from defection-dominating to cooperation-dominating ($b/N - c = 0$).  \\ \\

@@

\begin{align}
\pi_i(\D, \abf_{-i})&= 
\begin{cases}
\displaystyle \sum_{i=1}^N b_i \alpha(a_i) \cdot \frac{1}{N \cdot \mathcal{C}(\D,\abf_{-i})} \cdot \left(1 + w + w^2 + ...w^{\mathcal{C}(\D,\abf_{-i}) - 1} \right) &\quad  \text{if } \mathcal{C}(\D,\abf_{-i}) \neq 0 \\[15pt]
0 &\quad  \text{if } \mathcal{C}(\D,\abf_{-i}) = 0
\end{cases} \\[15pt]
\pi_i(\C, \abf_{-i}) &= \displaystyle \sum_{i=1}^N b_i \alpha(a_i) \cdot \frac{1}{N \cdot \mathcal{C}(\C,\abf_{-i})} \cdot \left(1 + w + w^2 + ...w^{\mathcal{C}(\C,\abf_{-i}) - 1} \right) - c_i
\label{Eq:payoff-synergistic-asymmetric}
\end{align} \\
\noindent where $\mathcal{C}(a,\abf_{-i})$ counts the number of cooperators in state $(a,\abf_{-i})$ and $\alpha(.)$ maps the actions $\C$ and $\D$ to 1 and 0 respectively. Note that the number of cooperators in the two states are related as: $\mathcal{C}(\D,\abf_{-i}) = \mathcal{C}(\C,\abf_{-i}) - 1$. We are interested in studying the long term stationary behaviour of players in this game when they learn through introspection. We first discuss results from the symmetric public goods game and then discuss results for the game with asymmetric players.\\

@@

In all the examples that we have studied so far, players can only choose between two actions (pure strategies). Introspection dynamics is particularly useful when players can use larger strategy sets. In this section, we study the stationary behaviour of players in a game where each player has 16 possible pure strategies. To this end, we adopt the multiplayer cooperation and rewarding game from Pal and Hilbe \cite{Pal:NatCom:2022}. In this game, there are two stages: in stage 1, players decide whether or not they contribute to a linear public good and in stage 2, they decide whether or not they reward their peers. When a player contributes to the public good, they pay a cost $c_i$ but generate a benefit worth $r_i c_i$ that is equally shared by everyone. When a player rewards a peer, they provide them a benefit of $\rho$ while incurring the cost of rewarding, $\gamma_i$ to self. In between the stages, players get full information about the contribution of their peers. In the rewarding stage, players have four possible strategies: they can either reward all the peers who contributed (social rewarding), reward all the peers who defected (antisocial rewarding), reward all peers irrespective of contribution (always rewarding) or reward none of the peers (never rewarding). Before stage 1 commences, player $i$ knows with some probability, $\lambda_i$, the rewarding strategy of all their peers. In stage 1, players can have four possible strategies: they can either contribute or defect unconditionally or they can be conditional cooperators or conditional defectors. Conditional cooperators (or defectors) contribute (or do not contribute) when they have no information about their peers (which happens with probability $1 - \lambda_i$). When a conditional player, $i$, knows the rewarding strategy of all their peers (which happens with probability $\lambda_i$) and finds that there are $n_{\mathrm{SR}}$ social rewarders and $n_{\mathrm{AR}}$ antisocial rewarders among his peers, they cooperate if and only if the marginal gain from rewards for choosing cooperation over defection outweighs the effective cost of cooperation. That is,

@@

\section*{Discussions and conclusion}

@@

We introduce introspection dynamics in $N$-player (a)symmetric games. In this learning model, at each time, one of the $N$ players updates (or not) their strategy by comparing the payoffs of two strategies only: the one being currently played and a random prospective one. Clearly, this assumption implies a simple cognitive process. Players do not optimize over the entire set of strategies as, for example, in best-responde models \cite{Gaunersdorfer:GEB:1995, Hofbauer:GEB:2005}. Furthermore, although conceptually similar, our model is also simpler than typical reinforcement learning models. For example, while we only have selection strength as a parameter (apart from payoffs), in Macy and Flache (2002) \cite{Macy:PNAS:2002}, there is a learning rate parameter (which could be comparable to our selection strength) but also an aspiration parameter which sets a payoff reference. In our model, the payoff reference is always the current one. All in all, whereas at each single time step, individuals are restricted to reason over two strategies only, as they iterate this step over time, they are able to fully explore the whole set of strategies, in a trial-and-error fashion. \\

\noindent Importantly, our model is also much simpler than the stochastic evolutionary game theory framework. While they both can involve solving the stationary distribution of a Markov process, they differ greatly in the state space size. Population models typically assume individuals play multiple games against (potentially all) other players in a population. As such, the state is defined by the number of players playing each strategy in the population(s). The number of states rapidly increases with the population size, the number of strategies, of players and of different types (in the case of asymmetric games). One can see how the mathematical analysis of multiplayer asymmetric games can become cumbersome. To deal with this issue, previous models frequently resorted to additional approximations, like low mutation rate \cite{Fudenberg:JET:2006, Veller:JET:2016} and weak selection \cite{Wild:JTB:2007}. %Most previous evolutionary models focus on limiting regimes, namely low mutation rate and selection strength approximations, as they allow for analytical results. 
On the contrary, in introspection dynamics, the states of the Markov process correspond to the outcome of a single (focal) game: for a $N$-player game, where player $i$ has $m_i$ possible actions, there are $m_1 \times m_2 \times ... \times m_N$ states. This feature hugely reduces our state space size, which is key for obtaining exact results.  \\

>>>>

\begin{document}
\maketitle
\onehalfspacing
\section*{Abstract}

@@

Most of these situations cannot be captured by the sum of several pairwise interactions. Thus, to account for such non-linearities, one needs to consider multiplayer games \cite{Gokhale:DGAA:2014}. For example, well-known effect that only emerges when more than two players are present is the ``second-order free-riding problem" \cite{Fowler:PNAS:2005}. A natural solution to maintain pro-social behavior in a community is to monitor and punish defectors (and/or reward cooperators). However, most forms of sanctioning are considerably costly \cite{Henrich:Science:2006}. Therefore, an additional (second-order) dilemma 
%is at stake% 
emerges: individuals would like cooperation to be incentivized but they prefer that others pay the associated costs.
%deterring defection is beneficial for all, but an individual that does not pay the associated cost has an advantage over the ones that do. As long as cooperation is incentivized by punishment (or rewards) of others, individuals can take advantage by defecting in this second-order dilemma. 
%As such, this tempts everyone not to engage in costly punishment (or rewarding), and cooperation can break down. For this reason, peer-punishment and sanctioning institutions can be hard to sustain without additional mechanisms \cite{Panchanathan:Nature:2004, Perc:SciRep:2012, Hilbe:SciRep:2012, Couto:JTB:2020, Pal:NatCom:2022}.

@@

Another interesting effect that can be explored with multiplayer games is the scale or size of the interaction itself. In situations that require some sort of coordination and where expectations on others play an important role in one's decisions, a growing group size might hinder the optimal outcome \cite{Skyrms:book:2003}. Likewise, it has been shown that it is hard to cooperate in large groups \cite{Santos:PNAS:2011, Hilbe:JTB:2015}. This is however, not a general effect in any interaction \cite{gokhale:JTB:2011}. 
%This is not a general effect, though. It can happen that adding more players actually promotes cooperation. Gokhale and Traulsen (2011) investigate the 2-player and 3-player cases of a task allocation problem \cite{gokhale:JTB:2011}. Although the underlying rules defining the interactions are the same for the two cases, simply introducing a third player changes the overall game, resulting in a decrease in free-riding.

@@

Additionally, 
%interaction% 
group size can vary in a population of players. There, not only the average group size can have an important effect, but also the variance of the group size distribution \cite{Pena:Evolution:2011, Broom:BMB:2019}.

@@

\noindent Complexity further increases when players differ significantly among themselves. This diversity can be captured by asymmetric games \cite{Taylor:JAP:1979, Schuster:AB:1981, Gaunersdorfer:TPB:1991, Hofbauer:JMB:1996, Hofbauer:GEB:2005, Ohtsuki:JTB:2010, McAvoy:PlosCB:2015, Veller:JET:2016, Hauser:Nature:2019}. In symmetric games, all players are indistinguishable. Thus, to fully characterise the state of the game, we only require to know the number of players playing each strategy. Conversely, in asymmetric games, players can differ in their available actions and in their incentives to choose each action.
%;players are of different types or roles. 
Therefore, they can have uneven effects on others' payoffs too. For example, in public goods games and collective-risk dilemmas, players can have different initial endowments (or wealth), productivitites, costs, risk perceptions, or risk exposures \cite{Milinski:CC:2011, Vasconcelos:PNAS:2014, Abouchakra:JTB:2014, Hauser:Nature:2019, Merhej:JAIR:2022, Wang:PTRSB:2023}. Hence, to fully describe the state of the game, we need to know the exact action of each player. This greatly increases the size of the game's state space; even more so, for more than two players. \\
%Hence, to fully describe the state of the game, we need to know which of the players is doing what. This greatly increases the size of the game's state space; even more so, for more than two players.  \\ 

@@

% deterministic EGT
\noindent Models from evolutionary game theory (EGT) \cite{Maynard-Smith:Nature:1973, Maynard-Smith:book:1982, Hofbauer:book:1998, Nowak:book:2006} and learning theory \cite{Sandholm:BioSys:1996, Fudenberg:book:1998b, Macy:PNAS:2002, Pangallo:GEB:2022}, have been widely used to study strategic behavior. 
%From evolutionary game theory (EGT) \cite{Maynard-Smith:Nature:1973, Maynard-Smith:book:1982, Hofbauer:book:1998, Nowak:book:2006} to learning models \cite{Sandholm:BioSys:1996, Fudenberg:book:1998b, Macy:PNAS:2002, Pangallo:GEB:2022}, game theoretic frameworks have been widely used to study strategic behavior. %, in particular, in multiplayer and asymmetric games.
%In evolutionary game theory, we assume a population of players and a mechanism by which strategies evolve over time. More successful strategies tend to persist in the population, either by genetic or cultural inheritance. The first key concept in evolutionary game theory is ``evolutionary stable strategy" (ESS),  introduced in 1973, in the seminal paper of John Maynard Smith and George Price \cite{Maynard-Smith:Nature:1973}. A (monomorphic) population playing an ESS cannot be invaded by mutants. 
\noindent The concept of evolutionary stable strategy, originally proposed for parwise encounters \cite{Maynard-Smith:Nature:1973}, was extended to multiplayer games \cite{Palm:JMB:1984, Broom:BMB:1997, Bukowski:IJGT:2004}. Also the well-known replicator equation \cite{Taylor:MB:1978, Hofbauer:book:1998} 
%easily comprises 
can easily account for multiplayer games \cite{Hauert:JTB:2006a, gokhale:PNAS:2010, Pena:Evolution:2011, Cressman:PNAS:2014, Pena:JTB:2014}. More recently, the replicator-mutator equation was applied to study the dynamics of multiplayer games, too \cite{Duong:DGAA:2020}. 
As for asymmetric games, a few additional assumptions are needed in the description of the model. For example, if there are two different types of players, typically, either there are two populations co-evolving (``bimatrix games" \cite{Hofbauer:book:1998, Gokhale:PRSB:2012, Tuyls:SciRep:2018}) or there is a single population of players where each can play the two types or roles (``role games") \cite{Hofbauer:book:1998}. 
%For role games, a strategy must include the action played in each of the two roles. 
The case of asymmetric games with more than two players is substantially less studied within deterministic EGT. Gokhale and Traulsen (2012) and Zhang et al. (2022) are two exceptions \cite{Gokhale:PRSB:2012, Zhang:arxiv:2022}. Notably, although these works study multiplayer games, they consider, at most, two different types (drawn from two populations), which leaves out the exploration of full asymmetry.
% models of multiple populations seem unnatural
% stochastic EGT
Also stochastic evolutionary game dynamics \cite{Nowak:Nature:2004, Traulsen:bookchapter:2009} provides several models for studying multiplayer and asymmetric games. 
%Here, the quantities of interest are fixation probabilities and fixation times (i.e., the probability and the average time that a single mutant overtakes an initially monomorphic population) for when mutations are rare \cite{Fudenberg:TPB:2006}, or the average strategy abundance (i.e., the stationary distribution of strategy frequencies) for when mutation rates take intermediate values \cite{antal:JTB:2009a, antal:JTB:2009b}. 
%These frameworks have been applied to asymmetric games assuming co-evolving (finite) subpopulations of different player types. 
Fixation probabilities \cite{Fudenberg:TPB:2006} in asymmetric 2-player \cite{Sekiguchi:DGA:2017}, asymmetric 3-player games \cite{Sekiguchi:DGAA:2022}, and symmetric multiplayer games \cite{Kurokawa:PRSB:2009, gokhale:PNAS:2010} were recently derived. Furthermore, average strategy abundances \cite{antal:JTB:2009a, antal:JTB:2009b} were obtained only for 2-player asymmetric games \cite{Ohtsuki:JTB:2010, Sekiguchi:PA:2013} or multiplayer symmetric games \cite{gokhale:JTB:2011, Wu:Games:2013, Kroumi:JMB:2022}. For a review on evolutionary multiplayer games both in infinitely large populations as well as in finite populations, we refer to Gokhale and Traulsen (2014) \cite{Gokhale:DGAA:2014}. 
Learning models (of strategic behavior) take a different approach from EGT \cite{Sandholm:BioSys:1996, Fudenberg:book:1998b, Macy:PNAS:2002, Hofbauer:GEB:2005, Tuyls:bookchapter:2005, Galla:PNAS:2013, Barfuss:PRE:2019, Barfuss:PNAS:2020, Pangallo:GEB:2022}. There is no evolution of strategies in a population necessarily, but a process by which individuals learn strategies dynamically. \\

@@

%[Learning models]
%learning models in general
%Despite the conceptual disparities, there have been already some efforts in unifying the two lines of research of evolutionary game theory and multiagent reinforcement learning \cite{Macy:PNAS:2002, Tuyls:bookchapter:2005, Bloembergen:JAIR:2015, Zhang:arxiv:2022}. 
% Clearly, we can distinguish two different pathways that led to ``learning in games": one which stemmed from the broad field of reinforcement learning \cite{Sandholm:BioSys:1996} and the other from the classical game theory itself \cite{Fudenberg:book:1998b}. The ``theory of learning in games" was proposed as a refinement of classical equilibrium concepts, like Nash equilibrium, where strict assumptions of common knowledge of rationality are made. Instead, learning consists of ``\textit{an alternative explanation that equilibrium arises as the long-run outcome of a process in which less than fully rational players grope for optimality over time}" \cite{Fudenberg:book:1998b}. In this context, several learning models were introduced, such as fictitious play \cite{Gaunersdorfer:GEB:1995} or perturbed best response dynamics \cite{Hofbauer:GEB:2005}.
%learning models for asymmetric and multiplayer games? fictious play for multiplayer games \cite{fudenberg:book:1998b} 
%can we say most of learning models have no analytical or exact results?
\noindent Introspection dynamics has recently proven to be a useful learning model for tackling asymmetric games \cite{Couto:NJP:2022, Hauser:Nature:2019, McAvoy:PNASnexus:2022, Schmid:PlosCB:2022}. In here, players update their strategies by exploring their own set of strategies in the following simple way: each time, after a round of the game, a random player considers a random alternative strategy; they compare the payoff that it would have given them to their current payoff; if the new strategy would provide a higher payoff, it is more likely adopted on the next round. We describe the model formally in the next section. 
While in Couto et al. (2022) \cite{Couto:NJP:2022} only 2-player games were considered, this framework is general enough to account for multiple players.  Particularly, introspection dynamics allows a natural exploration of full asymmetry in many-player games compared to population models. For example, in imitation dynamics, one needs to specify who is being imitated by whom \cite{Vasconcelos:PNAS:2014}. When players differ, it might not make sense to assume that they imitate others. Introspection avoids this assumption because players' decisions only depend on their own payoffs.
\\ \\
%[Our proposal]
\noindent Here, we extend previous results \cite{Couto:NJP:2022} of pairwise games under introspection dynamics to multiplayer games. 
First, we derive a formula that allows us to numerically compute the stationary distribution of introspection dynamics for any multiplayer asymmetric game. 
Second, we obtain explicit expressions of the stationary distribution for two special cases. These cases are additive games (where the payoff difference that a player gains by unilaterally switching to a different action is independent of the actions of their co-players), and symmetric multiplayer games with two strategies. 
To illustrate our theoretical results, we analyse various multiplayer asymmetric social dilemmas, extending the framework in \cite{Hauert:JTB:2006a} to asymmetric games. Finally, we also study the asymmetric version of a public goods game with a rewarding stage \cite{Pal:NatCom:2022}.

@@

\noindent We use the Eq. (\ref{Eq:stationary_dist_lpgg}) to analyze the stationary behavior of players after a long run of introspection in a linear public goods game (LPGG). First, we study the simplest case where all players are symmetric (the cost and benefit for all the 4 players are $c$ and $b$). Since all players are identical, the states of the game can be enumerated by counting the number of cooperators in the state. There can only be 5 distinct states of the game (from 0 to 4 cooperators).

@@

\noindent Naturally, $\beta$ plays an important role in determining the overall cooperation in the long run. When $\beta$ is low, average cooperation varies weakly with the strength of the dilemma, $b/N - c$ (Fig. \ref{Fig:LPGG-symmetric}c). Even when the temptation to defect is high ($b/N - c = -2$), players cooperate with a non-zero probability. Similarly, when cooperation is highly beneficial and strictly dominates defection ($b/N - c = 2$), players defect sometimes. At higher values of $\beta$, the stationary behavior of players is more responsive to the payoffs and thus reflects an abrupt change near the parameters where the game transitions from defection-dominating to cooperation-dominating ($b/N - c = 0$).  \\ \\

@@

\begin{align}
\pi_i(\D, \abf_{-i})&= 
\begin{cases}
\displaystyle \sum_{i=1}^N b_i \alpha(a_i) \cdot \frac{1}{N \cdot \mathcal{C}(\D,\abf_{-i})} \cdot \left(1 + w + w^2 + ...w^{\mathcal{C}(\D,\abf_{-i}) - 1} \right) &\quad  \text{if } \mathcal{C}(\D,\abf_{-i}) \neq 0 \\[15pt]
0 &\quad  \text{if } \mathcal{C}(\D,\abf_{-i}) = 0
\end{cases} \\[15pt]
\pi_i(\C, \abf_{-i}) &= \displaystyle \sum_{i=1}^N b_i \alpha(a_i) \cdot \frac{1}{N \cdot \mathcal{C}(\C,\abf_{-i})} \cdot \left(1 + w + w^2 + ...w^{\mathcal{C}(\C,\abf_{-i}) - 1} \right) - c_i
\label{Eq:payoff-synergistic-asymmetric}
\end{align} \\
\noindent where $\mathcal{C}(a,\abf_{-i})$ counts the number of cooperators in state $(a,\abf_{-i})$ and $\alpha(.)$ maps the actions $\C$ and $\D$ to 1 and 0 respectively. Note that the number of cooperators in the two states are related as: $\mathcal{C}(\D,\abf_{-i}) = \mathcal{C}(\C,\abf_{-i}) - 1$. We are interested in studying the long term stationary behavior of players in this game when they learn through introspection. We first discuss results from the symmetric public goods game and then discuss results for the game with asymmetric players.\\

@@

In all the examples that we have studied so far, players can only choose between two actions (pure strategies). Introspection dynamics is particularly useful when players can use larger strategy sets. In this section, we study the stationary behavior of players in a game where each player has 16 possible pure strategies. To this end, we adopt the multiplayer cooperation and rewarding game from Pal and Hilbe \cite{Pal:NatCom:2022}. In this game, there are two stages: in stage 1, players decide whether or not they contribute to a linear public good and in stage 2, they decide whether or not they reward their peers. When a player contributes to the public good, they pay a cost $c_i$ but generate a benefit worth $r_i c_i$ that is equally shared by everyone. When a player rewards a peer, they provide them a benefit of $\rho$ while incurring the cost of rewarding, $\gamma_i$ to self. In between the stages, players get full information about the contribution of their peers. In the rewarding stage, players have four possible strategies: they can either reward all the peers who contributed (social rewarding), reward all the peers who defected (antisocial rewarding), reward all peers irrespective of contribution (always rewarding) or reward none of the peers (never rewarding). Before stage 1 commences, player $i$ knows with some probability, $\lambda_i$, the rewarding strategy of all their peers. In stage 1, players can have four possible strategies: they can either contribute or defect unconditionally or they can be conditional cooperators or conditional defectors. Conditional cooperators (or defectors) contribute (or do not contribute) when they have no information about their peers (which happens with probability $1 - \lambda_i$). When a conditional player, $i$, knows the rewarding strategy of all their peers (which happens with probability $\lambda_i$) and finds that there are $n_{\mathrm{SR}}$ social rewarders and $n_{\mathrm{AR}}$ antisocial rewarders among his peers, they cooperate if and only if the marginal gain from rewards for choosing cooperation over defection outweighs the effective cost of cooperation. That is,

@@

\section*{Discussion and conclusion}

@@

We introduce introspection dynamics in $N$-player (a)symmetric games. In this learning model, at each time, one of the $N$ players updates (or not) their strategy by comparing the payoffs of two strategies only: the one being currently played and a random prospective one. Clearly, this assumption implies a simple cognitive process. Players do not optimize over the entire set of strategies as, for example, in best-responde models \cite{Gaunersdorfer:GEB:1995, Hofbauer:GEB:2005}. Furthermore, although conceptually similar, our model is also simpler than typical reinforcement learning models. For example, while we only have selection strength as a parameter (apart from payoffs), in Macy and Flache (2002) \cite{Macy:PNAS:2002}, there is a learning rate parameter (which could be comparable to our selection strength) but also an aspiration parameter which sets a payoff reference. In our model, the payoff reference is always the current one. All in all, while at each single time step individuals are restricted to reason over two strategies only, as they iterate this step over time, they are able to fully explore the whole set of strategies, in a trial-and-error fashion. \\

\noindent Importantly, our model is also much simpler computationally than the stochastic evolutionary game theory framework. While they both can involve solving the stationary distribution of a Markov process, they differ greatly in the state space size. Population models typically assume individuals play multiple games against (potentially all) other players in a population. As such, the state is defined by the number of players playing each strategy in the population(s). The number of states rapidly increases with the population size, the number of strategies, size of interaction and types of players (in the case of asymmetric games). One can see how the mathematical analysis of multiplayer asymmetric games can become cumbersome. To deal with this issue, previous models frequently resorted to additional approximations, like low mutation rate \cite{Fudenberg:JET:2006, Veller:JET:2016} and weak selection \cite{Wild:JTB:2007}. %Most previous evolutionary models focus on limiting regimes, namely low mutation rate and selection strength approximations, as they allow for analytical results. 
On the contrary, in introspection dynamics, the states of the Markov process correspond to the outcome of a single (focal) game: for a $N$-player game, where player $i$ has $m_i$ possible actions, there are $m_1 \times m_2 \times ... \times m_N$ states. This feature hugely reduces our state space size, which is key for obtaining exact results.  \\
